Seagrass mapping and monitoring \\ along the coasts of Crete, Greece

### 2011-02-02 MScDiplom_Lemenkova.tex
finished 'List of Acronyms', added hyperreferences, logos, footnotes

\documentclass[11pt, a4paper]{article}
\usepackage{color}
\usepackage{setspace}
\usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
\usepackage[subfigure]{tocloft}
%\usepackage[table]{xcolor}
\usepackage{paralist}
\usepackage{makeidx}
\makeindex
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}
\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}

@@

\begin{titlepage}
	\centering
	\title{Seagrass mapping and monitoring along the coasts of Crete, Greece}
	\author{by Lemenkova Polina}
	\date{\today}
	\maketitle	
\end{titlepage}
\pagebreak

@@

\begin{itemize}
	\item[*] University of Southampton (UK)\\
	\item[*] Lund University (Sweden)\\
	\item[*] University of Warsaw (Poland)\\
	\item[*] University of Twente, Faculty ITC (The Netherlands)
\end{itemize}
\pagebreak

@@

\textbf{Thesis Assessment Board}:
\begin{itemize}
	\item Name Examiner 1 (Chair)\dotfill
	\item Name Examiner 2\dotfill
	\item External Examiner \dotfill \\ \hfill{Dr. \textit{Petter Pilesj\"o} (Lund University, GIS Centre)}
\end{itemize}

@@

This research was carried out at the University of Twente, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)}, Enschede, the Netherlands, under the financial support of the EU \href{http://ec.europa.eu/education/external-relation-programmes/doc72_en.htm}{Erasmus Mundus Scholarship}. 
\vspace{1mm}\\
I thank my scientific supervisors, Dr. \textit{Albertus G.Toxopeus}  and Mr. \textit{Valentijn Venus} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work. \\
I also express them my most sincere appreciations for our fieldwork on Crete. \\ Besides scientific routine, we had an unforgettable time spent together. \\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g. SCUBA gear diving, instruments for underwater measurements, etc). I also appreciate their very warm, personal attitude and interest towards our research. 
\vspace{1mm}\\
My sincere gratitude goes again to Mr.\textit{Valentijn Venus} for his generous providing me with a lot of spatial data   relevant to my research (e.g. such as digital vector layers of Crete island, raster georeferenced maps for the whole island for my GIS project, Excel spreadsheets, etc). Without these vital data the implementation of this project would have hardly been possible. I am also thankful to him for the technical help in Google Earth imagery capturing, using GDAL technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc). \\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
I am thankful to Dr.rer.nat.\textit{Peter Gege}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work. I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career. He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth. \\ My special thank for his help with Python scripting. \\
I am also heartfelt thankful to him for his personal friendship and support. 
\vspace{1mm}\\
I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery. 
\vspace{1mm}\\
I thank Dr. Ir. \textit{Cornelis A.J.M. de Bie} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010. I used this skills on Crete. 
\vspace{1mm}\\
Many thanks for Dr.\textit{Dingtian Yang} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.

@@

The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones. The only flowering plant in the world that is able to live completely submerged, seagrasses play vital role in the marine ecosystems of the World Ocean. Seagrasses are the most important component in the environmental food chain of the coastal ecosystems, being a vital food source for various marine species (e.g. fish, dugongs, turtles, swans), and a producer of organic matter, which is the very basis of the food web.\\
The \textit{P.oceanica} seagrass is an endemic for the Mediterranean region, and a main species in the marine coastal environment of Greece. Meadows of \textit{P.oceanica} are subjected to the human activities, because they occur in coastal areas, where they are affected both by anthropogenic and by climatic and environmental factors.  Nowadays \textit{P.oceanica} is in the alarming state of regression, because of the deterioration of the environment in the Mediterranean Sea. Due to these reasons, \textit{P.oceanica} is a protected species since 1988 in some European countries (France). Monitoring \textit{P.oceanica} is therefore an important contribution to the saving and protecting the environment of Mediterranean region.\\
The current MSc thesis focuses on the monitoring of seagrass \textit{P.oceanica} along the northern coasts of Crete Island, Greece, and investigates the application of the remote sensing techniques for the seagrass mapping. 
This research was articulated in two parts, where the first one involves an ecological approach to the seagrass distribution in various regions around the globe and the experience of seagrass monitoring nowadays. The second part of this work has technical character and investigates the application of the remote sensing techniques towards seagrass mapping. It, furthermore, focuses on the optical properties of the \textit{P.oceanica} and other seafloor cover types, and studies distinguishability of various seafloor cover types. Studies of the optical characteristics of separate seafloor cover types were made with purpose to clarify, whether their spectral properties change with varying environmental conditions. \\
Special attention has been drawn on the role of environmental factors on the distribution of \textit{P.oceanica} along the coasts of Crete, and in particular, how the optical properties of the seafloor cover types, i.e. spectral reflectance, are being changed under varying external conditions, e.g. water column, amount of suspended particles and sediments in the seawater, and water temperature. For this purpose we studied differences in the spectral reflectance of \textit{P.oceanica} and other bottom cover types at three distinct depths. The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions. We applied WASI simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (MERIS, SeaWiFS, CZCS and MODIS), in order to understand their suitability and possible limitations for the seagrass mapping.\\
Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos). The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer. Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.

@@

\section*{Document Outline}
	\markright{Structure of the MSc Thesis}
	\addcontentsline{toc}{section}{Document Outline}
\textit{Chapter 1}: \nameref{sec:1} is an introductory section. It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work. In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.
\newline
In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment. We considered not only the Mediterranean environment, but also papers from Australian and Chinese scientists, because seagrass monitoring is most actively developed in the southern regions of our planet. The review of the existing RTM algorithms for the retrieval of the optical parameters, as well as description of various tools for the spectrometric measurements - are given in the same Chapter as well. More close attention has been given to the imaging hyperspectral radiometer Trios RAMSES, used for the data collection.\\
\textit{Chapter 3}: \nameref{sec:3} deals with the methods and materials used in current work. We start our discussion from the fieldwork area location and describe tools and instruments used during the fieldwork, as well as sampling design. Procedures of the data pre-processing and capturing imagery from the Google Earth are also presented in the same chapter. \\
\textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data. It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely WASI colour simulation and plotted resulting graphs. We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass. Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter. This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:6}.\\
A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}. Due to the standard editorial limitations of the current MSc work, it was not possible to include all them to the main chapters. However, we tried to make the structure of the Appendices most clear and easy-readable as possible, by dividing the Appendices into 10 various sub-sections and referencing to them from the main text where necessary. \\
The \nameref{bib} section includes literature and internet resources used for the current work. The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.\\
An Index is placed in the end of the Document to help the reader find what he is looking for.

@@

\section*{List of Acronyms}
	\markright{Acronyms}
	\addcontentsline{toc}{section}{List of Acronyms}
\doublespacing
\textbf{4AOP}: Automatized Atmospheric Absorption Atlas Operational Release \\
\textbf{6SV1}:  Second Simulation of a Satellite Signal in the Solar Spectrum, Version1\\
\textbf{ANOVA}: ANalysis of VAriance\\
\textbf{ASTER}: Advanced Spaceborne Thermal Emission and Reflection Radiometer\\
\textbf{BRDF}:  Bidirectional Reflectance Distribution Function\\
\textbf{CASI}:		Compact Airborne Spectrographic Imager\\
\textbf{CNES}: 		Centre National d'Etudes Spatiales (of France)\\
\textbf{CRTM}:		Community Radiative Transfer Model\\
\textbf{CZCS}:		 Coastal Zone Color Scanner\\
\textbf{DISORT}:	 Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium\\
\textbf{ENVISAT}:	ENVIronmental SATellite\\
\textbf{ESA}:		European Space Agency\\
\textbf{ERS}: 		European Remote Sensing satellite\\
\textbf{GIS}:		Geographic Information System\\
\textbf{GNU}:		GNU's Not Unix \\
\textbf{GPS}:		 Global Positioning System\\
\textbf{GRASS}:	 Geographic Resources Analysis Support System\\
\textbf{GRETL}:		GNU Regression, Econometrics and Time-series Library\\
\textbf{HyMap}:	Hyperspectral Mapper\\
\textbf{HyperOCR}:	Hyperspectral Ocean Colour Radiometer\\
\textbf{IKONOS}: 	derived its name from the Greek term \textit{eikon} for image\\
\textbf{ILWIS}:		Integrated Land and Water Information System\\
\textbf{iPAQ}:	internet Pocket PC by CompAQ\\
\textbf{KOPRA}:	 Karlsruhe Optimized and Precise Radiative transfer Algorithm\\
\textbf{LaTeX}:		Lamport Tau Epsilon Chi (Greek for \textit{techne}), modified by L.Lamport\\
\textbf{LAI}:	 Leaf Area Index\\
\textbf{Landsat TM}:	 Landsat Thematic Mapper \\
\textbf{Landsat ETM+}:	Landsat Enhanced Thematic Mapper Plus\\
\textbf{LyX}: 		LyriX-files\\
\textbf{MERIS}:		MEdium Resolution Imaging Spectrometer\\
\textbf{MIPAS}:		Michelson Interferometer for Passive Atmospheric Sounding\\
\textbf{MODIS}:	Moderate Resolution Imaging Spectroradiometer\\
\textbf{NASA}:		National Aeronautics and Space Administration\\
\textbf{NIR}:		 Near-Infra-Red\\
\textbf{NOAA}:		 National Oceanic and Atmospheric Administration\\
\textbf{OBC}:		On Board Calibration\\
\textbf{Olympus ST}: 	 Olympus Stylus Tough\\
\textbf{Open EV}: 	Open Earth View\\
\textbf{RAMSES}:	RAdiance Measurement SEnSor\\
\textbf{RRTM}:		Rapid Radiative Transfer Model\\
\textbf{RTTOV}: 	Radiative Transfer for TIROS Operational Vertical Sounder\\
\textbf{RTM}:		Radiative Transfer Model\\
\textbf{SAM}:		Scanning Acoustic Microscope\\
\textbf{SCUBA}:	Self Contained Underwater Breathing Apparatus\\
\textbf{SeaWiFS}: 	Sea-viewing Wide Field-of-view Sensor\\
\textbf{SPOT}:		Satellite Pour l'Observation de la Terre\\
\textbf{SPSS	}:	Statistical Package for the Social Sciences\\
\textbf{STDEV}:		Standard Deviation\\
\textbf{SWIR}: 		Short Wave InfraRed\\
\textbf{TIR}: 		Thermal InfraRed\\
\textbf{TIROS}: 	Television and InfraRed Observation Satellite\\
\textbf{UML}:		Unified Modelling Language\\
\textbf{UV}:		UltraViolet (10 nm to 400 nm)\\
\textbf{UVA}:		UltraViolet A, or long waves (315 nm-400 nm)\\
\textbf{UVB}:		 UltraViolet B, or medium waves (280 nm-315 nm)\\
\textbf{VIS}:		VISible\\
\textbf{VNIR}:		Visible and Near-InfraRed\\
\textbf{WASI}:		WAter colour SImulator
\pagebreak
\singlespace
\section*{List of Tools}\label{sec:0}
	\markright{List of Tools}
	\addcontentsline{toc}{section}{List of Tools}\index{Software used}

@@

reports the application of the CZCS images towards the case study of the Italian coast and shows

@@

For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}}\index{Seawater optical properties!RTM!WASI} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.

>>>>

\documentclass[11pt, a4paper]{article}
\usepackage{authblk}
%\usepackage{authorindex}
\usepackage{color}
\usepackage[stable]{footmisc }
\usepackage{setspace}
\usepackage[printonlyused,withpage]{acronym}
\usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
\usepackage[subfigure]{tocloft}
%\usepackage[table]{xcolor}
\usepackage{paralist}
\usepackage{makeidx}
\makeindex
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}
\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}

@@

\begin{titlepage}
	\centering
	\title{Seagrass mapping and monitoring \\ along the coasts of Crete, Greece}
	\author[1]{by Lemenkova Polina}
	\affil[1]{\textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
	\date{\today}
	\maketitle	
\end{titlepage}
\pagebreak

@@

\begin{itemize}
	\item[*] University of Southampton (UK)\\
	\item[*] Lund University (Sweden)\\
	\item[*] University of Warsaw (Poland)\\
	\item[*] University of Twente, Faculty ITC (The Netherlands)
\end{itemize}

\begin{figure}[H]
	\centering
	\subfloat{\includegraphics [width=0.2\textwidth] {Logo-Sou.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-Lun.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-War.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.5\textwidth] {ITC-logo.jpg}}	
\end{figure}

@@

\textbf{Thesis Assessment Board}:
\begin{itemize}
	\item Name Examiner 1 (Chair)\dotfill
	\item Name Examiner 2\dotfill
	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}}(Lund University, GIS Centre)}
\end{itemize}

@@

This research was carried out at the University of Twente, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)}, Enschede, the Netherlands, under the financial support of the EU \href{http://ec.europa.eu/education/external-relation-programmes/doc72_en.htm}{Erasmus Mundus Scholarship}. 
\vspace{1mm}\\
I thank my scientific supervisors, Dr. \textit{Albertus G.Toxopeus}  and Mr. \textit{Valentijn Venus} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work. \\
I also express them my most sincere appreciations for our fieldwork on Crete. \\ Besides scientific routine, we had an unforgettable time spent together. \\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g. SCUBA gear diving, instruments for underwater measurements, etc). I also appreciate their very warm, personal attitude and interest towards our research. 
\vspace{1mm}\\
My sincere gratitude goes again to Mr.\textit{Valentijn Venus} for his generous providing me with a lot of spatial data   relevant to my research (e.g. such as digital vector layers of Crete island, raster georeferenced maps for the whole island for my GIS project, Excel spreadsheets, etc). Without these vital data the implementation of this project would have hardly been possible. I am also thankful to him for the technical help in Google Earth imagery capturing, using GDAL technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc). \\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
I am thankful to Dr.rer.nat.\textit{Peter Gege}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work. I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career. He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth. \\ My special thank for his help with Python scripting. \\
I am also heartfelt thankful to him for his personal friendship and support. 
\vspace{1mm}\\
I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery. 
\vspace{1mm}\\
I thank Dr. Ir. \textit{Cornelis A.J.M. de Bie} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010. I used this skills on Crete. 
\vspace{1mm}\\
Many thanks for Dr. \author{\textit{Dingtian Yang}} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.

@@

The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones. The only flowering plant in the world that is able to live completely submerged, seagrasses play vital role in the marine ecosystems of the World Ocean. Seagrasses are the most important component in the environmental food chain of the coastal ecosystems, being a vital food source for various marine species (e.g. fish, dugongs, turtles, swans), and a producer of organic matter, which is the very basis of the food web.\\
The \textit{P.oceanica} seagrass is an endemic for the Mediterranean region, and a main species in the marine coastal environment of Greece. Meadows of \textit{P.oceanica} are subjected to the human activities, because they occur in coastal areas, where they are affected both by anthropogenic and by climatic and environmental factors. Nowadays \textit{P.oceanica} is in the alarming state of regression, because of the deterioration of the environment in the Mediterranean Sea. Due to these reasons, \textit{P.oceanica} is a protected species since 1988 in some European countries (France). Monitoring \textit{P.oceanica} is therefore an important contribution to the saving and protecting the environment of Mediterranean region.\\
The current MSc thesis focuses on the monitoring of seagrass \textit{P.oceanica} along the northern coasts of Crete Island, Greece, and investigates the application of the remote sensing techniques for the seagrass mapping. 
This research was articulated in two parts, where the first one involves an ecological approach to the seagrass distribution in various regions around the globe and the experience of seagrass monitoring nowadays. The second part of this work has technical character and investigates the application of the remote sensing techniques towards seagrass mapping. It, furthermore, focuses on the optical properties of the \textit{P.oceanica} and other seafloor cover types, and studies distinguishability of various seafloor cover types. Studies of the optical characteristics of separate seafloor cover types were made with purpose to clarify, whether their spectral properties change with varying environmental conditions. \\
Special attention has been drawn on the role of environmental factors on the distribution of \textit{P.oceanica} along the coasts of Crete, and in particular, how the optical properties of the seafloor cover types, i.e. spectral reflectance, are being changed under varying external conditions, e.g. water column, amount of suspended particles and sediments in the seawater, and water temperature. For this purpose we studied differences in the spectral reflectance of \textit{P.oceanica} and other bottom cover types at three distinct depths. The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions. We applied \ac{WASI} simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (\ac{MERIS}, \ac{SeaWiFS}, \ac{CZCS} and \ac{MODIS}), in order to understand their suitability and possible limitations for the seagrass mapping.\\
Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos). The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer. Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.

@@

\section*{Document Outline}
	\markright{Structure of the MSc Thesis}
	\addcontentsline{toc}{section}{Document Outline}
\textit{Chapter 1}: \nameref{sec:1} is an introductory section. It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work. In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.
\newline
In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment. We considered not only the Mediterranean environment, but also papers from Australian and Chinese scientists, because seagrass monitoring is most actively developed in the southern regions of our planet. The review of the existing RTM algorithms for the retrieval of the optical parameters, as well as description of various tools for the spectrometric measurements - are given in the same Chapter as well. More close attention has been given to the imaging hyperspectral radiometer Trios RAMSES, used for the data collection.\\
\textit{Chapter 3}: \nameref{sec:3} deals with the methods and materials used in current work. We start our discussion from the fieldwork area location and describe tools and instruments used during the fieldwork, as well as sampling design. Procedures of the data pre-processing and capturing imagery from the Google Earth are also presented in the same chapter. \\
\textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data. It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely \ac{WASI} colour simulation and plotted resulting graphs. We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass. Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter. This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:6}.\\
A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}. Due to the standard editorial limitations of the current MSc work, it was not possible to include all them to the main chapters. However, we tried to make the structure of the Appendices most clear and easy-readable as possible, by dividing the Appendices into 10 various sub-sections and referencing to them from the main text where necessary. \\
The \nameref{bib} section includes literature and internet resources used for the current work. The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.\\
An Index is placed in the end of the Document to help the reader find what he is looking for.

@@

\section*{List of Acronyms\footnote{The page is given where the glossary is first entered and defined}}
	\markright{Acronyms}
	\addcontentsline{toc}{section}{List of Acronyms}
%\singlespacing
\begin{acronym}
	\acro{4AOP}{\href{http://www.noveltis.fr/4AOP/interface/index.php}{Automatized Atmospheric Absorption Atlas Operational Release}} 
	\acro{6SV1}{Second Simulation of a Satellite Signal in the Solar Spectrum Version1} 
	\acro{ASCII}{\href{http://en.wikipedia.org/wiki/ASCII}{American Standard Code for Information Interchange}}
	\acro{ANOVA}{\href{http://en.wikipedia.org/wiki/Anova}{ANalysis of VAriance}} 
	\acro{ASTER}{\href{http://asterweb.jpl.nasa.gov/}{Advanced Spaceborne Thermal Emission and Reflection Radiometer}} 
	%\acro{BRDF}{Bidirectional Reflectance Distribution Function}
	\acro{CASI}{\href{http://www.borstad.com/casi.html}{Compact Airborne Spectrographic Imager}}
	\acro{CNES}{\href{http://www.cnes.fr/}{Centre National d'Etudes Spatiales (of France)}}
	\acro{CRTM}{\href{http://www.jcsda.noaa.gov/projects_crtm.php}{Community Radiative Transfer Model}}
	\acro{CZCS}{\href{http://oceancolor.gsfc.nasa.gov/CZCS/}{Coastal Zone Color Scanner}}
	\acro{DISORT}{\href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium}}
	\acro{ENVISAT}{\href{http://envisat.esa.int/instruments/meris/}{ENVIronmental SATellite}}
	\acro{ESA}{\href{http://www.esa.int/}{European Space Agency}}
	%\acro{ERS}{European Remote Sensing satellite}
	\acro{GIS}{Geographic Information System}
	%\acro{GNU}{GNU's Not Unix}
	\acro{GloVis}{\href{http://glovis.usgs.gov/}{Global Visualization Viewer}}
	\acro{GPS}{Global Positioning System}
	\acro{GRASS}{\href{http://grass.fbk.eu/}{Geographic Resources Analysis Support System}}
	%\acro{GRETL}{GNU Regression, Econometrics and Time-series Library}
	\acro{HyMap}{\href{http://www.hyvista.com/}{Hyperspectral Mapper}}
	\acro{HyperOCR}{\href{http://www.satlantic.com/Hyperspectral_Radiometers/Hyperspectral_Radiometers.asp}{Hyperspectral Ocean Colour Radiometer}}
	\acro{ILWIS}{\href{http://www.ilwis.org/}{Integrated Land and Water Information System}}
	\acro{iPAQ}{\href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{Internet Pocket PC by CompAQ}}
	\acro{KOPRA}{\href{http://www.imk-asf.kit.edu/english/312.php}{Karlsruhe Optimized and Precise Radiative transfer Algorithm}}
	\acro{LAI}{\href{http://en.wikipedia.org/wiki/Leaf_Area_Index}{Leaf Area Index}}
	\acro{Landsat TM}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Thematic Mapper}}
	\acro{Landsat ETM+}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Enhanced Thematic Mapper Plus}}
	\acro{MERIS}{\href{http://envisat.esa.int/instruments/meris/}{MEdium Resolution Imaging Spectrometer}}
	\acro{MIPAS}{\href{http://envisat.esa.int/instruments/mipas/}{Michelson Interferometer for Passive Atmospheric Sounding}}
	\acro{MODIS}{\href{http://modis.gsfc.nasa.gov/}{Moderate Resolution Imaging Spectroradiometer}}
	\acro{NASA}{\href{http://www.nasa.gov/}{National Aeronautics and Space Administration}}
	\acro{NIR}{Near-Infra-Red}
	\acro{NOAA}{\href{http://www.osdpd.noaa.gov/ml/index.html}{National Oceanic and Atmospheric Administration}}
	\acro{OBC}{On Board Calibration}
	\acro{Olympus ST}{\href{http://www.olympusamerica.com/}{Olympus Stylus Tough}}
	\acro{RRTM}{\href{http://rtweb.aer.com/}{Rapid Radiative Transfer Model}}
	\acro{RTTOV-9}{\href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{Radiative Transfer for TIROS (Television and InfraRed Observation Satellite) Operational Vertical Sounder}}
	\acro{RTM}{Radiative Transfer Model}
	\acro{SAM}{Scanning Acoustic Microscope}
	\acro{SCUBA}{\href{http://www.padi.com/scuba/}{Self Contained Underwater Breathing Apparatus}}
	\acro{SeaWiFS}{\href{http://oceancolor.gsfc.nasa.gov/SeaWiFS/}{Sea-viewing Wide Field-of-view Sensor}}
	\acro{SPOT}{\href{http://www.spotimage.fr/}{Satellite Pour l'Observation de la Terre}}
	\acro{SPSS}{ \href{http://www.spss.com/}{Statistical Package for the Social Sciences}}
	\acro{STDEV}{Standard Deviation}
	\acro{SWIR}{Short Wave InfraRed}
	\acro{TIR}{Thermal InfraRed}
	\acro{Trios-RAMSES}{\href{http://trios.de/index.php?option=com_content&view=category&layout=trios&id=47&Itemid=76&lang=en}{RAdiance Measurement SEnSor}}
	\acro{UML}{Unified Modelling Language}
	\acro{USGS}{\href{http://www.usgs.gov/}{U.S. Geological Survey}}
	\acro{UV}{UltraViolet (10 nm to 400 nm)}
	\acro{UVA}{UltraViolet A, or long waves (315 nm-400 nm)}
	\acro{UVB}{UltraViolet B, or medium waves (280 nm-315 nm)}
	\acro{VIS}{VISible}
	\acro{VNIR}{Visible and Near-InfraRed}
	\acro{WASI}{\href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}}
\end{acronym}

\singlespace
\section*{List of Tools}\label{sec:0}
	\markright{List of Tools}
	\addcontentsline{toc}{section}{List of Tools}\index{Software used}

@@

reports the application of the \ac{CZCS} images towards the case study of the Italian coast and shows

@@

For our purpose we have chosen the \ac{WASI}\index{Seawater optical properties!RTM!WASI} \ac{RTM} software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
